Lab 2 Final Report: COVID-19 Case Rate vs Population Demographics and Mask Policy

### 2020-12-04 Jackson_Liu_Temlock_Zhang_Lab2_Draft_Report.Rmd
Finished regression table section and analysis

model3 <- lm(case_rate_100k ~ senior_per_100k 
              + poverty_per_100k 
              + unemployed_per_100k 
             + mask_public_bool
             + log(population_density), data = df)
coeftest(model3, vcov = vcovHC)
```

@@

When observing the table above, three elements stand out.

Firstly, there is a difference in the calculated standard error between the coefficient test run on the linear models and the regression table. Taking a look at the standard error for the poverty percentage coefficient, the value in the table returns a significant result with a standard error of 55.125, as opposed to the insignificant result of a standard error of 57.684. This follows for model 3, with the table showing a significant result with standard error of 50.872, as opposed to an insignificant result of a standard error of 59.583.

Secondly, following the first observation, when designating the poverty percentage variable as significant, the regression table suggests that both covariate variables (including the mask mandate variable) are indeed significant with the COVID case rate. For the poverty percentage effect, the coefficient highlights that there is a positive correlation between the poverty percentage and case rate (as poverty percentage increases, so does case rate). The practical significance of this shows that for every percentage increase of population below the federal poverty line, there is an increase of roughly 112 cases per 100k for model 2, and roughly 104 cases per 100k for model 3. As for the mask mandate variable, the coeeficient highlights a negative correlation between it and the case rate (if the mask mandate is implemented, the case rate decreases). The practical significance is such that if the mask mandate has been implemented, the case rate per 100k will decrease by roughly 1000 cases.

Thirdly, we notice that the R2 value increases by a large value for every additional variable that is included within the model. Initially, we start with a very low R2 value, suggesting that almost no variance is explained in model 1 and that the model is a poor fit and predictor for the data. However, although still low, there is close to a 9 times increase in the amount of variance that is explained in model 2. Finally, in model 3, the R2 value increases by roughly 2.75 times the previous value, and a decent amount of the variance can be seen to be explained by the model. The adjusted R2 that accounts for the increase in number of variable terms in the model also correspondingly increases as the covariates are added. It also holds a negative value for model 1, signifying that there is no variance that can be explained by the model, further highlighting the poor fit.

As a note, it should be stressed that the models and variables included in this regression table may change when the final report is created to provide a more descriptive analysis of the original research question.

>>>>

```{r}
model3 <- lm(case_rate_100k ~ senior_per_100k 
              + poverty_per_100k 
              + unemployed_per_100k 
             + mask_public_bool
             + log(population_density), data = df)
coeftest(model3)
```

@@

Initially addressing the progression of the models at a macro level, the regression table highlights the fact that as the models build upon one another, they seem to get better at explaining the relationship between the input and output variables. Between model 1, 2, and 3, both the R2 and stricter adjusted R2 value increases as variables are added. The R2 value reflects the amount of variation that can be explained in the model on a 0-1 scale, while the adjusted R2 value does the same but also takes into account the number of variables measured (as can be seen in the decreasing degrees of freedom, df). The adjusted R2 value jumps from 0.082 initially in model 1 to a value of 0.267 in model 2, which is more than a 200% increase. The value again increases, albeit by a smaller amount, between model 2 and a value of 0.376 in model 3, which represents a 40% jump. When the R2 value is not adjusted, this value increases to 0.439 for model 3. Although not extraordinarily high given the scale, the value shows that that model 3 still explains much of the variance and is a relatively good predictor of the data, especially when compared to model 1. An additional value that reflects this improvement is the fact that the F-statistic is significant for all three models. The F-statistic is used to test whether or not to use the full model (inclusive of all input variables) or a reduced one with fewer variables. Given that the null hypothesis for this test is that the reduced model should be over the full model, this can be safely rejected in favor of the full model.

When diving into each model more closely and investigating the individual variable coefficients themselves, there is further evidence for robustness in the results across the models.

In model 1, the coefficient of -0.174 for the senior rate per 100,000 is significant with a p-value of 0.0234 (per the regression results from the Model 1 section). When the covariate variables of unemployment rate and poverty rate are added in to model 2, the coefficient for senior rate adjusts to a greater magnitude of -0.231 while the p-value appreciably decreases to a value of 0.0015 (per the regression results from the Model 2 section). This explains that not only has the magnitude of effect of the senior rate on case rate increased, but that this increased effect is also more statistically significant. In addition, the coefficients of the covariate variables themselves are also highly significant with p-values of 0.0005 and 0.0059 for poverty rate and unemployment rate respectively. However, it should be noted that the magnitudes of these effects are very small, with poverty rate having a positive relationship of 0.002 per 100,000 with case rate while unemployment rate has a negative relationship of 0.005 per 100,000.

Moving onto model 3, the magnitude of the coefficient for senior rate increases once again to a value of -0.242 while the statistical significance increases with a decreased p-value of 0.0004 (per the regression results from the Model 3 section). As highlighted before, the more complex model does a better job of predicting the effect of the senior rate variable. Conversely, the p-values for the poverty rate and unemployment rate coefficients become less significant, as they increase to 0.002 and 0.049 respectively. However, there is almost no detectable change in the coefficient values, as poverty rate is still at 0.002 and unemployment rate increases to -0.004. The change in the p-values are likely a direct result of the added variables of mask policy and the logarithmic transform of population density, where these variables introduce some additional variance in the model 2 covariates. Notably, the coefficient for the mask policy covariate is significant with a p-value of 0.0027 and has by far the greatest effect on the outcome variable with a coefficient of -945.6. The populuation density covariate is not significant with a p-value of 0.3369.

In terms of practical significance, there is great discrepancy between the effect of the mask policy variable compared to the other three significant features. With a coefficient of -945.6, this means that given the binary mask policy variable, when a mask policy is implemented (value of 1), there is an decrease in case rate of roughly 946 per 100,000 compared to when the policy is not in place. The next largest effect is from the senior rate variable, where in the final model, the coefficient of -0.242 translates as a decrease of 0.242 cases per 100,000 with every additional senior per 100,000. More intuitively, it can be said that there is an decrease of roughly 1 case per 100,000 with every 5 additional seniors per 100,000. This effect is roughly 190 times smaller than the effect of wearing a mask. Similarly, the effects of both poverty rate and unemployment rate are nominal. For every 500 additional people per 100,000 who are below the poverty line, there is 1 additional case per 100,000, and for every 250 additional people per 100,000 who are unemployed, there is 1 fewer case per 100,000. In comparison with likewise minimal effect of the senior rate, these effects are 100 and 50 times smaller respectively.
